---
title: "India against COVID19"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(
    input_file,
    encoding=encoding,
    output_file=file.path(
      dirname(input_file), 
      out_dir, 
      'index.html'
      )
    )
  })
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
```

```{r setup libraries, include=FALSE}
  library(tidyverse)
  library(lubridate)
  library(viridis)
  library(readr)
  library(janitor)
  library(kableExtra)
  # library(plotly)
  source('./source_crowd_data.r')
  source('./source_official_data.r')
  source('./source_world_data.R')
```

```{r calculate data, include = FALSE}

  
  dc <- build_crowd_data()
  dc_i <- dc$dc_i
  dc_raw_i <- dc$dc_raw_i  
  dc_age_i <- dc_raw_i %>% 
    mutate(AgeBracket = as.numeric(AgeBracket)) %>% 
    filter(!is.na(AgeBracket)) %>% 
    ungroup()
  
  do_i <- build_official_data()
  d_i_wide <- rbind(do_i, dc_i)
  d_i_long <- d_i_wide %>% pivot_longer(
    -c(StateUt, Date, Source), 
    values_to = 'Cumulative', 
    names_to = 'Status')
  ## Statewise Data
  
  data_panindia_total<- d_i_wide %>% 
    group_by(Date, Source) %>% 
    summarize(Total = sum(Total)) %>%
    mutate(StateUt = "India Total") %>% 
    ungroup()

  data_states_total <- d_i_wide %>% 
    select(StateUt, Source, Date, Total) %>% 
    ungroup()
  
  data_i_total <- rbind(data_panindia_total, data_states_total)
  
  data_i_aggregated_long <- d_i_long %>%
    group_by(Date, Source, Status) %>% 
    summarise(Cumulative = sum(Cumulative)) %>% 
    filter( Date > dmy('03-03-2020')) %>%
    ungroup() 
    
  
  ## World Data
  from <- dmy('1-02-20')

  d_w_wide <- source_world_data()
  
  data_w_long <- d_w_wide %>% select(-c(Lat, Long)) %>% filter(Date > from) %>%
    pivot_longer(-c(StateUt, Country,  Date), values_to = 'Total', names_to = 'Status')
  
  data_total_w <- data_w_long %>%
    group_by(Date, Status) %>% 
    summarize( Total = sum(Total) )  %>% ungroup() 
  
  data_rate_w <- data_total_w %>% 
    group_by(Status) %>%
    mutate(TotalPrev = lag(Total, default = 0, order_by = Date )) %>% 
    mutate(DatePrev = lag(Date, order_by = Date )) %>%
    filter(!is.na(DatePrev)) %>% mutate(DelDay = as.integer(Date-DatePrev)) %>%
    mutate( Rate1 = (Total - TotalPrev)*100/(DelDay*Total) ) %>%
    ungroup() %>%
    filter(!(Status %in% c('Hospitalized', 'Deaths')))

```

<!-- ## Current Status -->


```{r echo=F, results='asis', layout="l-body-outset"}

  data_today <- data_i_aggregated_long %>%
    filter(Date == max(Date)) %>%
    pivot_wider(values_from = Cumulative, names_from = Source) %>%
    select(-c(Date)) 
  kable(data_today) %>% kable_styling(
    bootstrap_options = c("striped", "hover"), full_width = F, position = 'float_right')
  
```

**Crowd Sourced Data: ** https://www.covid19india.org/

**Official Data: ** Ministry of Health and Family Welfare Data hosted on this repository
https://github.com/rahulnyk/covid19_india_data

**World Data: ** Repository hosted by Johns Hopkins CSSE
https://github.com/CSSEGISandData/COVID-19

Last update at **`r Sys.time()`**


##  {.tabset .tabset-fade .tabset-pills}

### Pan India

####  Daily growth rate 

```{r echo=FALSE, fig.cap="Daily Growth rate of COVID19 cases in India", message=FALSE, warning=FALSE}

  data_i_rate <- d_i_wide  %>% select(StateUt, Date, Total, Source) %>% 
    group_by(Date, Source) %>%
    summarise(Total = sum(Total)) %>% group_by(Source) %>%
    mutate( TotalPrev = lag(Total, default = 0, order_by = Date)) %>%
    mutate(  DatePrev = lag(Date, order_by = Date) ) %>% 
    mutate(DateCurPrev = as.integer(Date - DatePrev) ) %>%
    filter( Date > ymd('2020-03-03')	) %>%
    mutate( DailyRate = round((Total - TotalPrev)*100/(DateCurPrev*TotalPrev) ) ) %>%
    ungroup()

  ggplot(
    data = data_i_rate,  
    aes(x = Date, y = DailyRate, group = Source, color = Source, fill=Source)
    ) + 
  geom_vline(xintercept = dmy('24-03-2020'), linetype = '11' ) + 
  annotate(
  "text", x = dmy('25-03-2020'), y = 34, 
  size = 5, label = "Lockdown", hjust = 0, fontface = 1 ) +
  theme_minimal() +
  geom_point(alpha = 0.3, size = 2) + geom_line(linetype = 1, alpha = 0.3) + 
  geom_smooth(alpha = 0.05, size=0.6) +
  labs(y = "Growth rate in Percentage")  + 
  ylim(c(-10, 50)) +
  theme(legend.position = c(0.2, 0.8) ) +
  scale_fill_viridis_d(end = 0.6) + scale_color_viridis_d(end=0.6)

```

#### Cumulative

```{r include = FALSE}
  data_label <- data_i_aggregated_long %>% filter(Status == 'Total') %>% 
    filter( Date > Sys.Date()-5 )
  data_split_status <- data_i_aggregated_long %>% filter(Status != 'Total') 
```


```{r echo=FALSE, fig.cap="Total number of COVID19 cases in India by Date", message=FALSE, warning=FALSE}
  
   

  ggplot(
      data = data_split_status,  
      aes(x = Date, y = Cumulative, fill=Status)
    ) +      
    theme_minimal() +
    geom_area(stat = 'identity') + facet_wrap(~Source, ncol = 1) + 
    geom_text(
      data = data_label, 
      aes(label = Cumulative), show.legend = F,
      angle = 0, vjust = 0.5, hjust = 1, nudge_y = 0, nudge_x = -0.5) + 
    geom_vline(xintercept = dmy('24-03-2020'), linetype = '11' ) + 
    annotate(
    "text", x = dmy('25-03-2020'), y = 2000, 
    size = 5, label = "Lockdown", hjust = 0, fontface =1 ) +
    theme(legend.position = c(0.2, 0.8)) +
    scale_fill_viridis_d(begin = 0.1, end = 1) + scale_color_viridis_d(begin = 0.1, end=1)

```

### Statewise 

#### Statewise growth as per crowd sourced data

```{r include = F }

  data_statewise <- data_i_total %>% group_by(Source) %>%
    mutate(Day = as.integer(strftime(Date, format = "%j")) ) %>%
    ungroup() %>%
    ###
    group_by(StateUt, Source) %>% 
    mutate( DaysSince0 = Day - min(Day)  ) %>% 
    mutate( MaxDays = max(DaysSince0) ) %>%
    mutate( MaxCases = max(Total) ) %>%
    ungroup() %>%
    ###
    mutate(label =  paste(MaxCases, StateUt, sep=" | ")) %>%
    filter(Date > dmy('01-03-2020'))


d_i_long_total <- d_i_long %>% 
  filter(Status == 'Total') %>% 
  filter(Date == max(Date)) %>%
  filter(Cumulative > 5) %>%
  filter(!(StateUt == 'unassigned'))

d_i_long_total$StateUt[d_i_long_total$StateUt == "Telengana"] <- "Telangana"


```


```{r echo=FALSE, fig.cap="Statewise Split", message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8}

  ggplot(
      data = data_statewise %>% filter(Source == 'Crowd Source'), 
      aes(x=Date, y=reorder(label, MaxCases), 
          fill = StateUt, size = Total^(0.3))
      ) +
    geom_point(shape = 21, color = 'white') + 
    geom_vline(xintercept = dmy('24-03-2020'), linetype = '11' ) + 
    annotate(
    "text", x = dmy('25-03-2020'), y = 34, 
    size = 4, label = "Lockdown", hjust = 0.5, vjust = -1, fontface =1, angle = 30  ) +
    theme_minimal() + 
    theme(
      legend.position = 'off',
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      # panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(),
      axis.text.y = element_text(size=10,face="bold"),
      axis.text.x = element_text(angle = 30, hjust=0)
    ) + 
    scale_color_viridis_d(option="inferno", end = 0.9) + 
    scale_fill_viridis_d(option="inferno", end = 0.9) + 
    scale_y_discrete(position = "right") +
    scale_x_date(position = "top", date_breaks = '7 days') +
    scale_radius(
      range = c(1, 15),
      trans = "identity",
      guide = "legend"
    ) + 
    coord_cartesian(clip = 'off') 

```

#### State wise split of cumulative number of cases as per last update.

```{r echo=FALSE, fig.cap="Statewise Split", message=FALSE, warning=FALSE, fig.height = 6, fig.width = 8}
  ggplot(
      data = d_i_long_total, 
      aes(x = reorder( StateUt, Cumulative ), y = Cumulative, fill = Source)
      ) + 
    geom_bar(position = 'dodge', stat = 'identity') +
    theme(
      legend.position = c(0.4, 0.8),
      panel.grid.major = element_blank(),
      panel.background = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
      ) + scale_fill_viridis_d(option = 'cividis', end = 0.9)
```

### Agewise 

#### Agewise split of cases in India

Data available for **`r nrow(dc_age_i)`** cases

```{r echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 3, fig.width = 8}

  ggplot(data = dc_age_i, aes(x = AgeBracket, fill=Status)) +
    geom_histogram(stat = "bin", position = "stack", bins = 40, color = 'white') +
    scale_fill_viridis_d(option = 'magma', end = 0.9) + 
    theme_minimal() + 
    theme(
      axis.text=element_text(size=8, color = "darkgrey"), 
      axis.title=element_text(size=10),
      panel.grid.major = element_blank(),
      plot.title = element_text(size = 12, hjust = 0),
      legend.title = element_blank(), 
      legend.position = c(0.8, 0.8)
    )

```

### World

#### Global Growth rate

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 8}
  pal <- "viridis"

  plot_theme <- theme_minimal() + theme(
    axis.text=element_text(size=8, color = "darkgrey"), 
    axis.title=element_text(size=10),
    axis.title.x = element_blank(),
    # panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    plot.title = element_text(size = 12, hjust = 0),
    legend.title = element_blank(), 
    legend.position = c(0.2, 0.8)
  )
  ggplot(data = data_rate_w,  aes(x = Date, y = Rate1, color = Status, fill = Status)) + 
    geom_smooth(aes(group = Status), alpha = 0.2, size = 0.6, linetype=1) + 
    geom_point(alpha = 0.2, size=2) + geom_line(alpha = 0.4) +
    labs(y = "Growth rate in Percentage") + # labs(title = "Groth rate by date") + 
    scale_fill_viridis_d(option=pal, end = 0.6) + 
    scale_color_viridis_d(option=pal, end = 0.6) + 
    ylim(c(-10, 50)) +
    plot_theme 

```

#### Global Cumulative cases

``` {r echo=FALSE,  message=FALSE, warning=FALSE, fig.height = 4, fig.width = 8}
  ggplot(
      data = data_total_w %>% filter(!(Status %in% c('Total'))), 
      aes(x = Date, y = Total/1000, fill = Status)
      ) + 
    geom_area(position = "stack", stat = "identity") +
    # scale_y_continuous(trans = 'log2') +
    labs(y = "Number of cases (thousands)") + # labs(title = "Reported number of cases by date") + 
    scale_fill_viridis_d(option=pal, begin = 0, end = 0.9) + 
    scale_color_viridis_d(option=pal, begin = 0, end = 0.9) +
    plot_theme
```